---
title: "Alternative to automatic globals detection"
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(furrr)
library(carrier)
```

## Introduction

When writing production code with furrr, you might not want to leave it up to automatic globals detection to ensure that everything is shipped off to the worker correctly. While it generally works well, there will always be edge cases with static code analysis that may cause it to either fail (very bad) or copy more objects over to the workers than is required (less bad, but still not great).

As an alternative, you can use the [carrier](https://CRAN.R-project.org/package=carrier) package by Lionel Henry to manually isolate all of the dependencies required by your function. This ensures that the function's environment contains _only_ what you have requested, and nothing more. Combine this with `furrr_options(globals = FALSE)` to turn off automatic globals detection, and you should end up with a production worthy way to use furrr.

## Crates

The idea behind carrier is to package up a function and all of its dependencies into a _crate_. A crate is just another function, but with a specialized function environment that carries around only the dependencies required for it to run. To create a crate, directly wrap a call to `function()` with `crate()`. 

```{r}
crt1 <- crate(function(x) {
  stats::var(x)
})

crt1(c(1, 5, 3))
```

You can also specify the function to crate using the anonymous function shorthand with `~`.

```{r}
crt1_anon <- crate(~stats::var(.x))

crt1_anon(c(1, 5, 3))
```

You must manually namespace all package function calls with `pkg::`, as the crated function runs in an environment where only the base package has been loaded.

```{r, error=TRUE}
crt2 <- crate(function(x) {
  var(x)
})

crt2(c(1, 5, 3))
```

Any "outside" dependencies that you require must also be manually specified when creating the crate, otherwise it won't be found.

```{r, error=TRUE}
constant <- 1.67

crt3 <- crate(~ .x + constant)

crt3(2:5)
```

You can specify these dependencies by either inlining them into the crate with `!!`, or by supplying them as name-value pairs to `crate()`:

```{r}
crt4 <- crate(function(x) {
  x + !!constant
})

crt4(2:5)
```

```{r}
crt5 <- crate(
  constant = constant,
  function(x) {
    x + constant
  }
)

crt5(2:5)
```

Crates have a nice print method, allowing you to see the size of the crate and what has been captured:

```{r}
crt5
```

One downside of inlining with `!!` is that it doesn't show up as a separate element in the print method:

```{r}
crt4
```

The last thing to know about crates is that it is generally required that you crate a new function. By "new", I mean that the function cannot have already been created beforehand and assigned a name, because this prevents `crate()` from assigning it the correct environment.

```{r, error=TRUE}
fn <- function(x) {
  x
}

crate(fn)
```

## Crates and furrr

Using a crated function with furrr is just like using any other function, except you won't need to rely on the automatic globals detection (unless you have globals in `.x` or `...` that need to be found, but this is somewhat rare).

As an example, when `crt5()` was created we also captured the `constant` dependency object in the function's environment. Since furrr serializes `.f` alongside its environment, the dependencies come along for free without the need to auto detect them.

```{r}
crt5
```

```{r}
plan(multisession, workers = 2)

opts <- furrr_options(globals = FALSE)

x <- list(1:10, 11:20)

future_map(x, crt5, .options = opts)
```

```{r, echo=FALSE}
plan(sequential)
```

If you crate a function with extra optional arguments, you can still pass those through using furrr:

```{r}
median_doubled <- crate(function(x, na.rm = FALSE) {
  stats::median(x, na.rm = na.rm) * 2
})
```

```{r}
plan(multisession, workers = 2)

opts <- furrr_options(globals = FALSE)

x <- list(c(1, NA, 2), c(4, 5, NA))

future_map_dbl(x, median_doubled, na.rm = TRUE, .options = opts)
```

```{r, echo=FALSE}
plan(sequential)
```

Crates are also a great way to avoid accidentally shipping unneeded dependencies. In the _Common Gotchas_ article, we discuss the following example of accidentally shipping this `big` object to each worker, even though it isn't required by the function itself. This function call should be extremely fast, but is significantly slower than expected because it has to serialize `big`.

```{r}
my_slow_fn <- function() {
  # Massive object - but we don't want it in `.f`
  big <- 1:1e8 + 0L
  
  future_map_int(1:5, ~.x)
}
```

```{r}
plan(multisession, workers = 2)

system.time({
  my_slow_fn()
})
```

```{r, echo=FALSE}
plan(sequential)
```

One way to avoid this is to use carrier to crate the function, isolating it from the surrounding environment.

```{r}
my_crated_fn <- function() {
  # Massive object - but we don't want it in `.f`
  big <- 1:1e8 + 0L
  
  fn <- crate(~.x)
  opts <- furrr_options(globals = FALSE)
  
  future_map_int(1:5, fn, .options = opts)
}
```

```{r}
plan(multisession, workers = 2)

system.time({
  my_crated_fn()
})
```

```{r, echo=FALSE}
plan(sequential)
```

